\subsection{Related Work}
\label{sec:relatedwork}

This section positions the thesis against current infrastructure for
intent-based execution, account delegation, agent payments, agent identity,
wallet policy enforcement, and blockchain agent platforms. The purpose is not
to rank these systems, but to identify the part of the state of the art that
they address and the part that remains open for Intent-Conditioned Value
Projection (ICVP). The central distinction is that most existing systems answer
questions about how an action can be expressed, authorised, routed, paid for, or
executed. ICVP addresses the earlier decision question: what should count as
value for a given asset or proposal under a stated intent, which evidence is
required for that interpretation, and how the resulting decision can be audited
before a value-relevant transaction is prepared.

\subsubsection{Intent Protocols and Solver Systems}

Intent-centric protocols replace low-level transaction construction with a
higher-level statement of desired outcome. In ERC-7683-style settlement, or in
solver-mediated systems such as CoW Protocol, UniswapX, and Across-style
cross-chain intents, the user expresses an outcome and specialised solvers
compete or coordinate to satisfy it \cite{ERC7683, Nag2025CoWMechanism,
UniswapXWhitepaper, AcrossCrosschainIntents}.

These systems are directly relevant because they show that user intent can be a
first-class input to blockchain execution. Their intent, however, is normally an
execution constraint: swap token $x$ for at least amount $y$, bridge funds to a
target chain, fill an order within a deadline, or satisfy a route under price
and slippage limits. The solver problem is therefore to find a valid execution
path for a specified outcome. It does not usually formalise how the outcome was
selected as valuable in the first place. For example, a solver can optimise a
swap route, but it does not decide whether holding, selling, collateralising, or
using the asset is the right interpretation of value under the user's stated
goal.

Relative to ICVP, intent protocols leave open the value-interpretation layer.
They can execute a chosen action efficiently, but they do not define an
intent-conditioned projection over evidence, do not instantiate Availability,
Durability, and Utility differently for each intent, and do not produce a
valuation trace explaining why the action is defensible under that intent.

\subsubsection{Account Abstraction and Delegated Execution}

Account abstraction generalises the account model so that authentication,
authorisation, fee payment, batching, and recovery can be programmed rather than
hard-coded into externally owned accounts. EIP-4337 introduces smart-account
execution through user operations and bundlers \cite{Buterin2021}. EIP-7702
allows an externally owned account to temporarily behave as if it had account
code for a transaction \cite{EIP7702}. Broader surveys of account abstraction
describe this movement as a shift from key-based accounts toward programmable
account semantics \cite{Wang2023, Duan2024}. Session-permission proposals such
as ERC-7715 extend this direction by letting users grant limited capabilities to
applications or agents \cite{ERC7715}.

This line of work is important for the thesis because it provides the technical
substrate for constrained delegation. A user can give an agent a session key, a
spending cap, a list of permitted contracts, or a limited execution window
without handing over the user's private key. These mechanisms improve wallet
safety and make agent-assisted execution more practical.

The limitation is that programmable authorisation is not the same as
intent-conditioned valuation. Account-abstraction mechanisms can check whether a
transaction is permitted under a policy, but they do not specify what evidence
the agent should have gathered before proposing that transaction. They also do
not require a valuation trace that records the stated intent, the evidence used,
the intent-specific meaning of each valuation dimension, the decision label, and
the human approval step. In the terminology of this thesis, account abstraction
can help enforce part of Trace-Gated Authorisation, but it does not by itself
provide Evidence-Completeness Safety or the projection $P_i$ and instantiation
$\Phi_i$ used by ICVP.

\subsubsection{Agent Payments and Agent Wallets}

Recent agent-payment systems focus on enabling software agents to pay for APIs,
services, and on-chain actions. Examples include x402-style HTTP payment flows,
AP2-style agent payment protocols, and developer toolkits such as Coinbase
AgentKit that combine wallet operations with agentic application logic
\cite{x402Protocol, AP2Spec, CoinbaseAgentKit}.

These systems address a practical bottleneck: agents need payment rails and
wallet interfaces if they are to act in open digital markets. They can let an
agent purchase data, pay for a service, interact with a smart contract, or
prepare a wallet operation on behalf of a user. This makes them complementary to
the execution side of the architecture proposed in Chapter~\ref{sec:design}.

Their scope is nevertheless different from ICVP. Payment rails and agent-wallet
SDKs define how an agent can hold, request, or transfer value. They do not define
what value means under a particular intent, which evidence is necessary before a
payment or transfer is recommended, or how a human reviewer can inspect a
structured valuation trace before authorising the action. They therefore
support machine payments and agent action, but not the contextual decision model
that this thesis studies.

\subsubsection{Agent Identity and Reputation}

Agent identity standards aim to make agents discoverable, describable, and
assessable as network participants. ERC-8004, for example, extends the NFT-based
agent model with metadata about agent capabilities and on-chain mechanisms for
ratings or reputation \cite{EIP8004}. This direction is relevant because an
agentic Internet of Value requires some way to know which agent acted, what it
claimed to be able to do, and how other participants assessed it.

Identity and reputation, however, answer a different question from ICVP. They
help identify or validate an agent, but they do not decide how a particular NFT,
productive asset, or treasury proposal should be valued under a specific user or
governance intent. A reputable agent may still use the wrong evidence, collapse
the decision to market price, or prepare a transaction without exposing a
complete trace. ERC-8004-style identity can therefore complement NFTAA-level
authorisation, but it does not replace the evidence projection and valuation
trace required by ICVP.

\subsubsection{Policy-Based Wallets and Custody}

Smart-contract wallets and custody systems provide policy enforcement for asset
movement. Safe modules and guards can restrict calls, require multisig approval,
or add custom checks before execution \cite{Liochon2023}. Institutional custody
systems such as Turnkey- or Fireblocks-style policy engines similarly enforce
approval workflows, spending limits, role-based controls, and destination
constraints \cite{TurnkeyPolicyEngine, FireblocksPolicyEngine,
ZetzscheNikolakopoulou2025CryptoCustody}.

This work is close to the enforcement side of the thesis. It shows that
transaction authorisation can be constrained by explicit policy rather than by a
single signing key. It also shows why policy engines are necessary when agents
prepare transactions that move assets or vote with governance rights.

The gap is that policy enforcement usually evaluates the transaction envelope:
who may sign, which destination is allowed, what amount may move, which contract
may be called, and which approvals are required. It does not usually evaluate
whether the transaction is supported by complete intent-required evidence, an
intent-instantiated $A/D/U$ analysis, and a human-approved valuation trace. In
this thesis, NFTAA uses policy as an enforcement layer, but the policy is
trace-gated: a value-relevant action should be authorised only when it is linked
to a valid trace, complete projected evidence, a \textit{proceed} decision, and
human approval.

\subsubsection{Treasury Governance Tooling}

On-chain treasury governance has its own tooling stack. Tally and Boardroom
aggregate proposals across DAO frameworks and provide voting interfaces.
Snapshot runs off-chain signalling votes that are later settled on-chain.
Polkassembly and Subsquare carry proposal discussion, sentiment, and voter
context for Polkadot OpenGov \cite{Valastin2024proto, PolkassemblyArchive,
SubsquareReferenda}. Karpatkey-style treasury dashboards report treasury
composition, exposure, and runway across DAOs.
Dune, Nansen, and Token Terminal expose on-chain spending, delegate
behaviour, and protocol-level metrics that voters consult during evaluation
\cite{TallyGovernanceDocs, BoardroomGovernanceAPI, SnapshotDocs, KpkDocs,
Saesen2026DAOGovernancePerformance}.

These tools matter for governance participation, and the thesis relies on the
evidence they expose. Their scope is presentation, aggregation, and voting
workflow, not intent-conditioned proposal valuation. They show voters which
proposals exist, what discussion they have attracted, and how they are
moving through governance. They do not project that evidence through a
stated stewardship intent, do not instantiate Availability, Durability, and
Utility for the proposal class, and do not emit an auditable valuation
trace that a deterministic validator can check before a vote is recommended.
They are complementary to ICVP, not a substitute for it.

\subsubsection{Valuation Approaches for NFTs, Productive Assets, and Proposals}

Asset-class-specific valuation work supplies many of the building blocks an
ICVP projection needs. NFT valuation has been studied through hedonic-pricing
models that combine traits, sales history, and rarity
\cite{Lohmer2024, Ante2021, Dowling2022, Nadini2021, Kong2021}, through
provenance- and metadata-based durability assessment, and through community
or access-rights signalling. Productive-asset valuation is dominated by
yield, peg, and protocol-risk analyses for liquid staking and DeFi positions,
covering audit history, validator-set composition, slashing exposure, and
liquidation-depth analysis \cite{frangella2022aave, gambacorta2024defi}.
Treasury-proposal valuation in academic and practitioner work mostly takes
the form of qualitative rubrics and case studies of voting behaviour
\cite{Rodrigues2022, Valastin2024proto}, sometimes paired with quadratic
funding that aggregates small contributions into a public-goods signal
\cite{Buterin2018QF, Buterin2019, Pasquini2021, Weyl2018}.

These approaches tell ICVP which evidence is informative inside a particular
regime, but they stay regime-specific. They do not share one decision
contract across asset classes, do not formalise how intent reshapes the
relevant evidence, and do not produce one valuation trace that policy
enforcement can gate on. ICVP reuses their domain insights inside the
regime-specific projections $P_i$ and instantiations $\Phi_i$. The
contribution at this layer is the cross-regime decision model and the trace
it emits, not a new asset-specific scoring formula.

\subsubsection{Blockchain Agent Platforms}

Existing blockchain agent platforms demonstrate concrete ways to combine
off-chain artificial intelligence with on-chain identity, wallets, incentives,
and execution. They are useful examples of the environment in which ICVP would
operate, but they should not be treated as the whole state of the art.

Virtuals Protocol enables the creation, co-ownership, and interaction with
tokenised AI agents \cite{VirtualsWhitepaper2024}. Each agent is represented as
an NFT with an associated token-bound account, so the agent can have a persistent
on-chain identity and wallet-like capabilities. The model is strongest as a
community-ownership and agent-tokenisation mechanism: economic upside and
governance can be shared among token holders. Its limitation for this thesis is
that value interpretation is mostly left to market dynamics, token incentives,
and platform-level governance. The platform does not formalise an
intent-conditioned evidence projection or require a valuation trace before an
agent-related asset action is authorised.

Autonolas, or Olas, provides decentralised infrastructure for autonomous
services that run off-chain while interacting with on-chain contracts
\cite{AICoin2023, OlasNetwork2024}. It separates agent code, service registries,
operators, and on-chain treasuries, and it is especially relevant for
fault-tolerant multi-agent operation. Autonolas therefore addresses service
coordination and operator incentives more directly than consumer wallet
delegation. It can support autonomous services that act across chains, but it
does not define the ICVP decision layer: the system does not require
intent-specific evidence completeness, intent-instantiated valuation dimensions,
or human-approved trace-gated transaction preparation for every value-relevant
action.

Griffain illustrates a user-facing natural-language agent experience on Solana.
The user states an operation, the agent parses the command, checks wallet state,
selects routes through services such as Jupiter, prepares and submits the
transaction, and accounts for usage through its energy model \cite{Griffain2024}.
This is valuable as a usability reference because it reduces blockchain
interaction to natural-language task delegation. The trade-off is that the
decision path is not an auditable ICVP-style valuation trace. A command can be
translated into an action, but the system does not formally expose which
evidence made the action valuable under the user's intent, how Availability,
Durability, and Utility were instantiated, or which missing-evidence condition
should force a \textit{defer} rather than a transaction proposal.

\begin{table}[!htb]
	\centering
	\caption{Related-work comparison against the ICVP thesis gap}
	\label{tab:related-work-icvp}
	\scriptsize
	\begin{tabular}{|p{3.0cm}|p{4.6cm}|p{5.7cm}|}
		\hline
		\textbf{Category} & \textbf{Main contribution} & \textbf{Limit relative to ICVP} \\
		\hline
		Intent protocols and solver systems & Express desired execution outcomes and route them through solvers. & Do not formalise intent-conditioned value interpretation, evidence projection, or $A/D/U$ instantiation. \\
		\hline
		Account abstraction and delegated execution & Provide programmable authorisation, session-like delegation, batching, recovery, and gas abstraction. & Authorise transactions, but do not require valuation traces or evidence-completeness checks. \\
		\hline
		Agent payments and agent wallets & Let agents hold wallets, pay for services, and initiate machine-to-machine payments. & Provide payment and action rails, but not a contextual value decision model. \\
		\hline
		Agent identity and reputation & Identify agents, describe capabilities, and attach reputation or ratings. & Validate who the agent is, not what value means under a stated intent. \\
		\hline
		Policy wallets and custody & Enforce spending limits, approvals, allowlists, and transaction policies. & Check policy compliance, but not trace-gated valuation backed by intent-required evidence. \\
		\hline
		Treasury governance tooling & Aggregate proposals, surface discussion and sentiment, and run voting workflow. & Present and aggregate proposals, but do not project them through stewardship intent into auditable valuation traces. \\
		\hline
		Valuation approaches per asset class & Provide regime-specific scoring, hedonic models, and quadratic-funding signals for NFTs, productive assets, and proposals. & Remain asset-class specific; do not share a cross-regime decision contract or emit policy-gateable valuation traces. \\
		\hline
		Blockchain agent platforms & Combine off-chain agents with on-chain identity, services, incentives, and user-facing execution. & Demonstrate useful agent architectures, but do not jointly provide ICVP projection, valuation traces, and NFTAA-level trace-gated authorisation. \\
		\hline
	\end{tabular}
\end{table}

\subsubsection{Resulting Gap}

The current state of the art provides many pieces required for agentic value
systems. Intent protocols and solver markets can execute specified outcomes.
Account abstraction and session delegation can make authorisation programmable.
Agent-payment systems can let machines pay and act. Agent identity and
reputation standards can describe and assess agents. Policy wallets and custody
systems can enforce transaction constraints. Treasury governance tools aggregate
proposals and run voting workflow. Asset-class-specific valuation work supplies
regime-level evidence. Platforms such as Virtuals, Autonolas, and Griffain show
that AI agents can be connected to on-chain identity, services, incentives, and
user interfaces.

Existing systems give us execution, automation, settlement, custody, or
coordination primitives. They do not give us a formal model that projects
stated intent over heterogeneous evidence spaces into auditable valuation
traces and that preserves human approval and policy-constrained execution.
That is the joint decision-and-authorisation layer studied in this thesis.
No existing work jointly provides intent-conditioned evidence projection,
intent-instantiated Availability, Durability, and Utility, auditable
valuation traces, human-approved transaction preparation, and NFTAA-level
trace-gated authorisation. ICVP addresses this gap by treating stated intent as
a projection over heterogeneous evidence spaces and by making the resulting
decision inspectable before any value-relevant action is authorised on-chain. The
claim is not that ICVP is objectively better than these systems. The claim is
that they mostly solve routing, delegation, identity, payment, or policy
enforcement problems, while ICVP targets the evidence-grounded interpretation of
value that must precede defensible agentic execution in the Internet of Value.
