%Analysis
\section{Internet of Value and Blockchain}
\label{sec:analysis}

When the Internet was created, it had a simple goal: connect computers so they could share data reliably, even if parts of the network failed.
As the technology matured the goal pivoted to enable people to communicate via computer and transfer data via the World Wide Web.
Anyone could spin up a simple Apache server at home and distribute HTML files over the web using the Hypertext Transfer Protocol (HTTP).
However, most pages were just static, stateless content. For most internet users, it was an unreal experience because they could enjoy features such as email and access to the latest news without waiting until the next day for a printed paper.
In 1996, when CSS1.0 was introduced in browsers, we could customize the web page's design to our preferences \cite{Choudhury2014}.

Furthermore, a few early e-commerce applications were created, such as Amazon.com.
This was an era called Web 1.0 or the Internet of Information. The web was a single-page application accessible from any browser.
Powered by speculation in new startups, Web 1.0 ended in the dot-com era \cite{Goodnight2010}.
A massive number of people joined in the era of Web 2.0. Although it did not bring any profound technological change, it was a new web era.

Web pages become more interactive thanks to the JavaScript language and frameworks such as jQuery. Web 2 mainly uses HTML5 and CSS3, and the primary fundamental is that the web has become a social and interconnected place.
What made Web 2.0 cool was the ability for anyone to create and share content online. This content could range from a blog post on WordPress to a YouTube video \cite{Web1, Krumm2008, Chakravorty2017}.
With the first release of the Facebook platform in 2004, the web became a place where spreading information is easier and faster.
These platforms allowed non-technical people to create blogs and websites.

Additionally, the Internet allowed the creation of shared economy. Instead of owning a car, we can rent a car and pay just for the time we use it \cite{Parker2016}.
As it is known, most vehicles are used only 5\% of the day \cite{Keith2022}. That means that the car is underutilized a lot. A similar situation is happening with housing properties.
Before the Internet, we rented a house or flat to friends or family members. The Internet allowed us to share our assets with strangers and lowered the transaction cost.
Platforms like Airbnb were created to make the renting process more accessible. We do not need to limit the gig economy only to asset sharing \cite{Cupi2022, Sundararajan2016}.
TaskRabbit allowed people to use their time to do tasks and earn money \cite{Prassl2015}.
The problem is that people participating in the gig economy must earn a reputation for each service separately.
For example, we can have an imposing reputation score in TaskRabbit, but if we want to become Airbnb hosts, we must build our reputation again \cite{Schor2016}.
This means we are looking for a new paradigm to solve this problem.

As described in Chapter~\ref{sec:blockchain}, blockchain technology provides the foundational infrastructure for decentralized, trustless value transfer through consensus-driven state transitions \cite{Nakamoto, wood2014ethereum}. Building on these primitives, the decentralized finance (DeFi) sector has demonstrated that traditional financial products, such as lending (Compound \cite{papangelou2023exploring}, AAVE \cite{gambacorta2024defi, frangella2022aave}), decentralized exchange (Uniswap \cite{Adams2021}), and asset management, can be replicated in a permissionless, composable manner \cite{Angeris2020, Kim2022}.

Without any doubt, the emergence of blockchain technology has enabled a fundamental shift from an Internet of Information to an Internet of Value. While the former revolutionized how we share and access data, the latter promises
to transform how we transfer, store, and exchange assets of any kind.
However, the term Internet of Value is often conflated with related but distinct concepts. ``Web3'' broadly refers to the decentralized web stack (identity, storage, computation), while ``decentralized finance'' denotes the specific application of blockchain to financial instruments. The Internet of Value is a broader paradigm: it encompasses not only financial transactions but any form of value transfer, including non-monetary assets such as reputation, access rights, intellectual property, and governance participation.

Lohmer et al. \cite{Lohmer2024} provide one of the most comprehensive definitions of the Internet of Value, describing it as follows:

\begin{quote}
The Internet of Value enables the direct transfer of assets of any kind as well as processing and settlement of business transactions based on blockchain technology in conjunction with digital platforms, tokens, smart contracts, and other technologies such as the Internet of Things and artificial intelligence. Its principles are to create a system compliant with data protection values, equal and fair access for all participants, effective representation of digital and physical assets, and disintermediation for direct peer-to-peer communication and the exchange of values. These technologies and principles facilitate open content and value sharing, transparency, sustainable behavior, and self-sovereign digital identities \cite{Lohmer2024}.
\end{quote}

From the definition, we can derive three essential elements that differentiate the Internet of Value from previous paradigms.
First, we treat blockchain as the foundation of the IoV rather than just a technology. To enable secure value transfer, we must ensure orchestration among multiple participants in the network, which leads us to the second element: establishing the Internet of Value as a fair and autonomous socio-economic framework rather than merely a technical infrastructure.

Third, and most importantly for this work, the definition implies that value itself is not a singular, objective quantity but a multi-dimensional construct. Through expert interviews and systematic analysis, Lohmer et al. \cite{Lohmer2024} later summarized this understanding to propose that value in the Internet of Value can be defined using three distinct dimensions: Economic worth, Utility or access rights, and Ethical and sustainability contributions.

What does this imply? It reveals that asset valuation cannot be based solely on price discovery. For instance, non-fungible tokens can have a low floor market price but possess high utility for a specific community \cite{Ante2021, Dowling2022, Nadini2021, Kong2021}. Similarly, fungible governance tokens can be cheap to obtain yet contribute meaningfully to on-chain decision-making processes. Traditional financial valuation frameworks, which prioritize market price and liquidity, fail to capture this complexity. The absence of such a multi-dimensional valuation framework, one that captures utility and contribution alongside price, is the first gap this thesis identifies and addresses in Chapter~\ref{sec:aims}.

What role does the blockchain play in autonomous agents?
Blockchain provides the infrastructure for storing and sharing data without the need for trusted intermediaries. Through tokenization, any asset can be represented on-chain with verifiable ownership and transferability \cite{Jeon2022}. This eliminates the risks inherent in centralized storage: when a company shuts down its servers, the assets stored there become inaccessible or worthless without further use \cite{Duan2021}.

By decentralizing these assets with a trustless consensus mechanism implemented on the blockchain \cite{Williams2019}, virtual assets will become tradeable across the network. They can be used as a store of value. For example, we can lock a rare virtual item from the game and acquire other tokens. As a bonus, this game item can be used as collateral for a loan \cite{Gudgeon2020}. The principle is straightforward. When we want to collateralize our asset, we need to find someone who can offer a specific amount for it. The second party will get the item if we do not repay the debt.
Centralized platforms that attempt to provide such features must introduce trust assumptions and single points of failure, contradicting the fundamental principle of the Internet of Value.

Furthermore, as the Internet of Value expands across multiple blockchains and incorporates an ever-growing diversity of asset
shapes from tokenized real-world assets and generative algorithms to blockchain-native protocols and services \cite{Bambridge2022}, and to intellectual property rights. This challenge extends to the funding of open-source projects and public goods \cite{Kukelka2022public, Calhoun2006, Ostrom1990}, where manual assessment for adequate funding becomes infeasible.
On the one hand, by focusing on primary open-source projects, we prevent double-spending during development time \cite{Sopko2022}. Moreover, having the possibility that multiple parties are contributing to the same codebase, the project is more stable, less prone to bugs, and more secure. Open-source code creates value. Additionally, non-technical people can contribute to the project by providing their ideas and suggestions \cite{Alves2022, Yin2021}.

Projects may seek funding through quadratic funding models or on-chain treasury proposals. In both cases, evaluating proposals requires processing extensive on-chain and off-chain data, which makes manual evaluation costly and inconsistent at scale. Agents are therefore better positioned as decision-support components that retrieve evidence, structure proposal information, and produce auditable recommendations for human voters rather than as autonomous deciders. Using the quadratic funding model, we are giving the project more value and having skin in the game \cite{Buterin2018QF, Buterin2019}. Other companies can then match the project's funding and move towards a greater goal \cite{Pasquini2021, Buterin2018, Freitas2021, Weyl2018}. The second path is to create a treasury proposal for on-chain voting. If the proposal passes, the project will get funding from the treasury \cite{Rodrigues2022, Valastin2024proto}. Treasury volume regularly exceeds what individual voters can review carefully. Polkadot's OpenGov system, for example, processes hundreds of proposals per year, each requiring analysis of on-chain spending history, team track records, deliverable feasibility, and alignment with ecosystem priorities \cite{Valastin2024proto}. Individual voters lack the time and tooling to perform this analysis consistently, which leads to voter apathy, low participation rates, and decisions driven by social influence rather than evidence. The framing this thesis adopts is that agents support such evaluation by retrieving evidence and structuring it into auditable recommendations. Final voting authority remains with human participants or formally authorised governance actors.
Agentic systems can serve as the interpretation layer that retrieves evidence, structures it across multiple dimensions, and proposes auditable recommendations \cite{Wang2024survey}. Blockchain remains the settlement layer that records approved actions in a transparent and immutable manner. The two layers are complementary rather than substitutable. The agent does not replace human judgement, and the chain does not interpret value.

However, several critical gaps prevent this synthesis from being realized in practice. First, despite growing literature on blockchain-based asset valorization \cite{DeLope2021}, no standardized framework exists for multi-dimensional, goal-oriented asset valuation.
Moreover, current account models do not cover the needs of autonomous agents. Externally owned accounts bind authority to a single key. Smart contract wallets, including recent account-abstraction proposals such as EIP-4337 \cite{Buterin2021, Wang2023, Duan2024}, focus on flexible authentication and gas sponsorship rather than agent-oriented delegation. None of these provides primitives for policy enforcement, risk isolation, and auditable decision-making that stay compatible with existing DeFi protocols, which limits what can safely be delegated to agents.
Last but not least, there is no established methodology for translating natural-language intent into verifiable on-chain actions while maintaining user sovereignty and transparency, with respect to the fact that user has to be aware of ongoing actions \cite{Amershi2019, Shneiderman2022, Wang2021}.


The identified gaps, namely the absence of a multidimensional value framework, the lack of agent-compatible on-chain primitives, and the scalability limitations of manual governance evaluation, motivate the research questions formulated in Chapter~\ref{sec:aims}.
