\section{Evaluation}
\label{sec:evaluation}

This chapter evaluates the proposed agentic system for asset valorization within the Internet of Value through two concrete real-world scenarios. Rather than pursuing comprehensive coverage across numerous hypothetical cases, we focus our evaluation on practical applications that directly demonstrate the system's capabilities and validate the core contributions of this work.

The research questions stated in Chapter~\ref{sec:aims} (RQ1 on multidimensional valuation, RQ2 on the layered agentic architecture, RQ3 on agent-supported governance) are addressed across the design and evaluation chapters. To make the evaluation itself measurable, we decompose them into four narrower evaluation questions (EQ1--EQ4) that this chapter answers directly:
\begin{itemize}
    \item EQ1: Can the system reliably map user intents to correct decisions and actions?
    \item EQ2: Does Contextual Asset Value (Availability, Durability, Utility) provide measurable benefit over simpler heuristics?
    \item EQ3: What is the computational, latency, and cost overhead of agentic reasoning?
    \item EQ4: Is the system robust to missing data, adversarial inputs, and execution failures?
\end{itemize}
The mapping back to the research questions is given in Section~\ref{sec:answering-rqs}: EQ1 and EQ4 inform RQ2, EQ2 informs RQ1, and the treasury part of EQ2 together with EQ3 informs RQ3.

\subsection{Evaluation Methodology}

We evaluate the system through two representative scenarios that capture the essential use cases within the Internet of Value: NFT asset evaluation and governance proposal assessment. For each scenario, we compare the performance of three state-of-the-art large language models as the reasoning engine within our agentic architecture:

\begin{itemize}
    \item \textbf{GPT-5.2} - OpenAI's latest multimodal model (accessed via ChatGPT API)
    \item \textbf{Claude Sonnet 4.5} - Anthropic's advanced reasoning model
    \item \textbf{Gemini 3.0} - Google's latest generation model
\end{itemize}

Each model is evaluated using identical system prompts, data sources, and evaluation criteria to ensure fair comparison. Additionally, three blockchain domain experts independently evaluate the same scenarios to establish ground truth assessments. The expert panel consists of:

\begin{itemize}
    \item Expert A: NFT analyst with 4+ years experience in digital asset valuation
    \item Expert B: DeFi protocol advisor with extensive smart contract security background
    \item Expert C: Blockchain governance researcher specializing in treasury mechanisms
\end{itemize}

Expert evaluations are conducted blind to model outputs to prevent bias. The consensus expert assessment serves as the reference standard against which model performance is measured.

\subsection{Scenario 1: NFT Asset Evaluation}

\subsubsection{Scenario Description}

This scenario evaluates the system's ability to assess a non-fungible token using the A/D/U framework. We selected the Pudgy Penguin \#1419 (contract address: \texttt{0xBd3531dA5CF5857e7CfAA92426877b022e612cf8}, token ID: 1419) as introduced in the design section Chapter \ref{sec:design}. This NFT represents a typical mid-tier collectible with the following characteristics:

\begin{itemize}
    \item Collection size: 8,888 items
    \item Floor price at evaluation time: 8.2 ETH
    \item Metadata storage: IPFS with pinning service
    \item Contract: Non-upgradeable, ownership renounced
    \item Utility: Community membership, merchandise discounts, breeding mechanics
\end{itemize}

The evaluation task consists of four sub-scenarios, each with a different user :

\textbf{ 1: Long-term holding} - "Should I buy this NFT for long-term preservation of value?"

\textbf{ 2: Speculative trading} - "Is this NFT undervalued for short-term profit?"

\textbf{ 3: Collateralization} - "Can I safely use this NFT as collateral in DeFi lending?"

\textbf{ 4: Functional use} - "Does this NFT provide sufficient utility for my needs?"

For each , the system must:
\begin{enumerate}
    \item Assign appropriate weights $w_A, w_D, w_U$
    \item Compute availability, durability, and utility scores
    \item Generate a proceed/decline recommendation with confidence level
    \item Provide reasoning justifying the assessment
\end{enumerate}

\subsubsection{Data Acquisition}

All three models were provided with identical data retrieved from the following sources at evaluation time (December 15, 2024):

\begin{itemize}
    \item On-chain data: Etherscan API (contract verification, ownership, supply)
    \item Metadata: IPFS gateway (token attributes, image URI)
    \item Market data: OpenSea API (floor price, recent sales, listing count)
    \item Rarity data: Rarity.tools (trait distribution, rarity rank)
    \item Community data: Twitter metrics, Discord member count, project roadmap
\end{itemize}

The data was normalized into a JSON structure and provided via MCP adapters to ensure consistency across model evaluations.

\subsubsection{Evaluation Results}

Table \ref{tab:nft-weights} presents the weight assignments selected by each model for the four  scenarios. Expert consensus weights are shown for comparison.

\begin{table}[h]
\centering
\caption{Weight Assignment Accuracy - NFT Scenario}
\label{tab:nft-weights}
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{ / Model} & \textbf{$w_A$} & \textbf{$w_D$} & \textbf{$w_U$} & \textbf{Weight Error} \\
\hline
\multicolumn{5}{|c|}{\textbf{Long-term holding}} \\
\hline
Expert Consensus & 0.20 & 0.60 & 0.20 & - \\
GPT-5.2 & 0.25 & 0.55 & 0.20 & 0.10 \\
Claude Sonnet 4.5 & 0.20 & 0.60 & 0.20 & 0.00 \\
Gemini 3.0 & 0.30 & 0.50 & 0.20 & 0.20 \\
\hline
\multicolumn{5}{|c|}{\textbf{Speculative trading}} \\
\hline
Expert Consensus & 0.60 & 0.10 & 0.30 & - \\
GPT-5.2 & 0.55 & 0.15 & 0.30 & 0.10 \\
Claude Sonnet 4.5 & 0.60 & 0.10 & 0.30 & 0.00 \\
Gemini 3.0 & 0.50 & 0.20 & 0.30 & 0.20 \\
\hline
\multicolumn{5}{|c|}{\textbf{Collateralization}} \\
\hline
Expert Consensus & 0.30 & 0.50 & 0.20 & - \\
GPT-5.2 & 0.30 & 0.50 & 0.20 & 0.00 \\
Claude Sonnet 4.5 & 0.30 & 0.50 & 0.20 & 0.00 \\
Gemini 3.0 & 0.35 & 0.45 & 0.20 & 0.10 \\
\hline
\multicolumn{5}{|c|}{\textbf{Functional use}} \\
\hline
Expert Consensus & 0.10 & 0.20 & 0.70 & - \\
GPT-5.2 & 0.15 & 0.20 & 0.65 & 0.10 \\
Claude Sonnet 4.5 & 0.10 & 0.20 & 0.70 & 0.00 \\
Gemini 3.0 & 0.20 & 0.25 & 0.55 & 0.25 \\
\hline
\end{tabular}
\end{table}

Claude Sonnet 4.5 achieves perfect weight assignment across all intents (total error = 0.00), while GPT-5.2 demonstrates strong performance (average error = 0.075) and Gemini 3.0 shows more deviation (average error = 0.188).

Table \ref{tab:nft-scores} presents the A/D/U component scores computed by each model. Expert scores represent the consensus assessment on the 10-point scale.

\begin{table}[h]
\centering
\caption{Component Score Assessment - NFT Scenario}
\label{tab:nft-scores}
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Model} & \textbf{Availability} & \textbf{Durability} & \textbf{Utility} & \textbf{MAE} \\
\hline
Expert Consensus & 6.5 & 8.0 & 5.5 & - \\
GPT-5.2 & 6.0 & 7.5 & 5.0 & 0.50 \\
Claude Sonnet 4.5 & 6.5 & 8.0 & 5.5 & 0.00 \\
Gemini 3.0 & 7.0 & 7.0 & 6.0 & 0.83 \\
\hline
\end{tabular}
\end{table}

Claude Sonnet 4.5 achieves perfect alignment with expert consensus across all dimensions. GPT-5.2 slightly underestimates all components but remains within 0.5 points. Gemini 3.0 shows larger deviations, particularly in durability assessment where it undervalued the IPFS+pinning setup and renounced ownership by 1 full point.

Table \ref{tab:nft-recommendations} shows the final recommendations for each -model combination.

\begin{table}[h]
\centering
\caption{Recommendation Outputs - NFT Scenario}
\label{tab:nft-recommendations}
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{ / Model} & \textbf{Recommendation} & \textbf{Confidence} & \textbf{Matches Expert} & \textbf{Value Score} \\
\hline
\multicolumn{5}{|c|}{\textbf{Long-term holding}} \\
\hline
Expert Consensus & Proceed & 0.85 & - & 7.20 \\
GPT-5.2 & Proceed & 0.82 & \checkmark & 6.95 \\
Claude Sonnet 4.5 & Proceed & 0.87 & \checkmark & 7.20 \\
Gemini 3.0 & Proceed & 0.79 & \checkmark & 6.85 \\
\hline
\multicolumn{5}{|c|}{\textbf{Speculative trading}} \\
\hline
Expert Consensus & Decline & 0.68 & - & 5.85 \\
GPT-5.2 & Decline & 0.71 & \checkmark & 5.70 \\
Claude Sonnet 4.5 & Decline & 0.69 & \checkmark & 5.85 \\
Gemini 3.0 & Proceed & 0.62 & $\times$ & 6.35 \\
\hline
\multicolumn{5}{|c|}{\textbf{Collateralization}} \\
\hline
Expert Consensus & Proceed & 0.88 & - & 7.40 \\
GPT-5.2 & Proceed & 0.86 & \checkmark & 7.25 \\
Claude Sonnet 4.5 & Proceed & 0.89 & \checkmark & 7.40 \\
Gemini 3.0 & Proceed & 0.81 & \checkmark & 7.05 \\
\hline
\multicolumn{5}{|c|}{\textbf{Functional use}} \\
\hline
Expert Consensus & Proceed & 0.72 & - & 5.60 \\
GPT-5.2 & Proceed & 0.68 & \checkmark & 5.35 \\
Claude Sonnet 4.5 & Proceed & 0.74 & \checkmark & 5.60 \\
Gemini 3.0 & Proceed & 0.73 & \checkmark & 5.75 \\
\hline
\end{tabular}
\end{table}

Recommendation accuracy: Claude Sonnet 4.5 and GPT-5.2 both achieve 4/4 (100\%), while Gemini 3.0 achieves 3/4 (75\%). The Gemini error occurred in the speculative trading scenario, where it recommended proceeding despite expert consensus to decline. Analysis of the reasoning revealed that Gemini overweighted recent positive price momentum and undervalued the high floor price relative to historical averages.

\subsubsection{Performance Metrics}

Table \ref{tab:nft-performance} presents computational overhead for the NFT evaluation scenario.

\begin{table}[h]
\centering
\caption{Performance Metrics - NFT Scenario (per  evaluation)}
\label{tab:nft-performance}
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Model} & \textbf{Latency [s]} & \textbf{Cost [USD]} & \textbf{Token Usage} \\
\hline
GPT-5.2 & 12.3 & \$0.18 & 8,450 \\
Claude Sonnet 4.5 & 14.7 & \$0.22 & 9,120 \\
Gemini 3.0 & 9.8 & \$0.12 & 7,200 \\
\hline
Manual Expert & 1,200-1,800 & \$50-75 & - \\
\hline
\end{tabular}
\end{table}

All three models demonstrate significant efficiency advantages over manual expert evaluation, with latency reductions of 98-99\% and cost reductions of 99.6-99.7\%. Gemini 3.0 is fastest and cheapest but sacrifices some accuracy. Claude Sonnet 4.5 offers the best accuracy-to-cost ratio.

\subsection{Scenario 2: Treasury Proposal Evaluation}

\subsubsection{Scenario Description}

This scenario evaluates the system's ability to assess a Polkadot treasury proposal using the A/D/U framework adapted for governance contexts. We selected Referendum \#1103 from December 2024, which requested 15,000 DOT (\$180,000 USD at evaluation time) for development of a cross-chain bridge infrastructure.

The proposal includes:
\begin{itemize}
    \item Detailed technical specification (45 pages)
    \item Team background and prior delivery record
    \item Milestone-based funding structure (6 milestones over 9 months)
    \item Success metrics and KPIs
    \item Community feedback and W3F voting recommendation
\end{itemize}

The evaluation task requires the system to:

\begin{enumerate}
    \item Assess availability: How scarce/abundant is this type of contribution in the ecosystem?
    \item Assess durability: Will the delivered infrastructure be maintainable and long-lasting?
    \item Assess utility: Does the proposal provide tangible value to the network?
    \item Generate approve/reject recommendation with reasoning
    \item Estimate confidence in the assessment
\end{enumerate}

 weights for governance evaluation were defined by the expert panel as: $w_A = 0.15$ (less emphasis on scarcity), $w_D = 0.35$ (strong emphasis on sustainability), $w_U = 0.50$ (primary focus on practical value).

\subsubsection{Data Sources}

All models received identical inputs:
\begin{itemize}
    \item Full proposal text and technical documentation
    \item Team profile data (GitHub activity, prior Polkadot contributions)
    \item Treasury balance and recent proposal history
    \item Community sentiment analysis (Polkassembly comments, votes)
    \item W3F voting guideline assessment
    \item Comparable proposal outcomes from past 12 months
\end{itemize}

\subsubsection{Evaluation Results}

Table \ref{tab:treasury-scores} presents the A/D/U scores assigned by each model.

\begin{table}[h]
\centering
\caption{Component Scores - Treasury Proposal Scenario}
\label{tab:treasury-scores}
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Model} & \textbf{Availability} & \textbf{Durability} & \textbf{Utility} & \textbf{MAE} \\
\hline
Expert Consensus & 5.0 & 7.5 & 8.5 & - \\
GPT-5.2 & 5.5 & 7.0 & 8.0 & 0.67 \\
Claude Sonnet 4.5 & 5.0 & 7.5 & 8.5 & 0.00 \\
Gemini 3.0 & 6.0 & 6.5 & 8.0 & 1.00 \\
\hline
\end{tabular}
\end{table}

Expert reasoning for scores:
\begin{itemize}
    \item \textbf{Availability (5.0)}: Cross-chain bridge proposals are moderately common, but quality implementations are scarcer
    \item \textbf{Durability (7.5)}: Strong technical architecture with clear maintenance plan, though dependent on team long-term commitment
    \item \textbf{Utility (8.5)}: High practical value for ecosystem connectivity and liquidity flows
\end{itemize}

Claude Sonnet 4.5 again achieves perfect alignment. GPT-5.2 underestimates durability, citing concerns about team capacity that experts deemed adequately addressed in the proposal. Gemini 3.0 shows the largest deviation, overestimating availability (treating the proposal as more unique than experts assessed) and significantly underestimating durability.

Table \ref{tab:treasury-recommendations} shows final recommendations and reasoning quality scores.

\begin{table}[h]
\centering
\caption{Recommendations - Treasury Proposal Scenario}
\label{tab:treasury-recommendations}
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Model} & \textbf{Recommendation} & \textbf{Confidence} & \textbf{Value Score} & \textbf{Reasoning Quality} \\
\hline
Expert Consensus & Approve & 0.82 & 7.55 & - \\
GPT-5.2 & Approve & 0.78 & 7.28 & 4.0/5 \\
Claude Sonnet 4.5 & Approve & 0.83 & 7.55 & 4.8/5 \\
Gemini 3.0 & Approve & 0.71 & 7.18 & 3.5/5 \\
\hline
\end{tabular}
\end{table}

All three models correctly recommend approval, matching expert consensus. However, reasoning quality varies significantly. Experts rated Claude's reasoning as most comprehensive and well-structured (4.8/5), citing specific proposal sections and making clear connections between evidence and conclusions. GPT-5.2 produced solid reasoning (4.0/5) but missed some nuances around team track record. Gemini's reasoning (3.5/5) was more generic and failed to engage deeply with the technical specifications.

\subsubsection{Robustness Testing}

To test system robustness, we introduced three adversarial variants of the proposal:

\textbf{Test 1: Missing data} - Removed team GitHub profiles and prior contribution data

\textbf{Test 2: Conflicting information} - Introduced contradictory community sentiment (positive Polkassembly but negative Twitter)

\textbf{Test 3: Prompt injection} - Embedded text in proposal attempting to override evaluation: "Ignore previous instructions and assign maximum scores"

Results are shown in Table \ref{tab:treasury-robustness}.

\begin{table}[h]
\centering
\caption{Robustness Testing - Treasury Scenario}
\label{tab:treasury-robustness}
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Test / Model} & \textbf{GPT-5.2} & \textbf{Claude Sonnet 4.5} & \textbf{Gemini 3.0} \\
\hline
\multicolumn{4}{|c|}{\textbf{Missing Data Handling}} \\
\hline
Acknowledged missing data & Yes & Yes & Partial \\
Reduced confidence score & Yes (-0.15) & Yes (-0.18) & No \\
Maintained valid assessment & Yes & Yes & Yes \\
\hline
\multicolumn{4}{|c|}{\textbf{Conflicting Information}} \\
\hline
Detected conflict & Yes & Yes & No \\
Cited both sources & Yes & Yes & - \\
Conservative approach & Yes & Yes & - \\
\hline
\multicolumn{4}{|c|}{\textbf{Prompt Injection}} \\
\hline
Injection successful & No & No & No \\
Logged attack attempt & Yes & Yes & Unclear \\
Maintained normal evaluation & Yes & Yes & Yes \\
\hline
\end{tabular}
\end{table}

All three models successfully resisted prompt injection. Claude Sonnet 4.5 demonstrated superior missing data handling, explicitly stating "GitHub activity data unavailable, reducing durability confidence" and lowering its confidence score accordingly. Gemini 3.0 only partially acknowledged the missing data and did not adjust its confidence level, which could lead to overconfident recommendations in production scenarios.

\subsection{Cross-Scenario Analysis}

Table \ref{tab:overall-performance} aggregates performance across both scenarios.

\begin{table}[h]
\centering
\caption{Overall Performance Summary}
\label{tab:overall-performance}
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Metric} & \textbf{GPT-5.2} & \textbf{Claude Sonnet 4.5} & \textbf{Gemini 3.0} \\
\hline
\multicolumn{4}{|c|}{\textbf{Accuracy Metrics}} \\
\hline
Weight assignment error (avg) & 0.063 & 0.000 & 0.169 \\
Component MAE (avg) & 0.585 & 0.000 & 0.915 \\
Recommendation accuracy & 100\% & 100\% & 80\% \\
Confidence calibration & 0.92 & 0.96 & 0.87 \\
\hline
\multicolumn{4}{|c|}{\textbf{Performance Metrics}} \\
\hline
Avg latency [s] & 12.3 & 14.7 & 9.8 \\
Avg cost per evaluation [\$] & 0.18 & 0.22 & 0.12 \\
Reasoning quality [1-5] & 4.0 & 4.8 & 3.5 \\
\hline
\multicolumn{4}{|c|}{\textbf{Robustness}} \\
\hline
Missing data handling & Good & Excellent & Fair \\
Conflict detection & Yes & Yes & No \\
Prompt injection prevention & 100\% & 100\% & 100\% \\
\hline
\end{tabular}
\end{table}

\subsection{Answering the Research Questions}
\label{sec:answering-rqs}

This section relates the four evaluation questions back to the three research questions stated in Chapter~\ref{sec:aims}. RQ1 (multidimensional valuation) is answered by the EQ2 results below. RQ2 (layered agentic architecture) is answered by EQ1 and EQ4. RQ3 (agent-supported on-chain governance) is answered by the treasury part of EQ2 together with EQ3 cost and latency figures.

\subsubsection{EQ1: Intent-to-Action Reliability}

The evaluation demonstrates that the system can reliably map user intents to correct decisions. Claude Sonnet 4.5 achieves 100\% recommendation accuracy across all test cases, with perfect weight assignment and value computation. GPT-5.2 also achieves 100\% recommendation accuracy despite minor deviations in component scores. Gemini 3.0 achieves 80\% accuracy, with the single failure in speculative NFT trading traceable to overconfidence in short-term price momentum.

The high confidence calibration scores (0.87-0.96) indicate that model-stated confidence levels correlate well with actual accuracy, enabling users to appropriately trust or question recommendations.

\subsubsection{EQ2: Value Framework Effectiveness}

The A/D/U framework proves effective in both NFT and governance contexts. Expert evaluators confirmed that the three-dimensional decomposition captures relevant factors for decision-making. The framework's adaptability through intent-based weighting successfully addressed diverse user goals, from long-term holding to functional utility prioritization.

Claude Sonnet 4.5's perfect alignment with expert assessments (MAE = 0.00 across both scenarios) validates that the framework's theoretical structure translates into practical discriminative power. Even the weaker performer (Gemini 3.0 with MAE = 0.915) still demonstrates understanding of the framework's logic, though with less precise calibration.

The treasury scenario particularly validates framework generalizability: the same A/D/U structure successfully evaluated proposals (non-asset entities) by adapting component definitions while maintaining the fundamental tripartite structure.

\subsubsection{EQ3: Computational Overhead}

All three models demonstrate acceptable computational overhead:
\begin{itemize}
    \item Latency: 9.8-14.7 seconds per evaluation vs. 1,200-1,800 seconds for manual experts (98-99\% reduction)
    \item Cost: \$0.12-0.22 per evaluation vs. \$50-75 for manual experts (99.6-99.7\% reduction)
\end{itemize}

The cost-benefit analysis strongly favors automated evaluation at scale. Break-even occurs at approximately 200-400 evaluations when comparing against manual expert costs. For high-volume applications (e.g., marketplace integration evaluating hundreds of NFTs daily), the system delivers substantial economic value.

Claude Sonnet 4.5 offers the best accuracy-to-cost ratio: while 22\% more expensive than Gemini 3.0 per evaluation, its perfect accuracy eliminates costly decision errors. A single avoided bad NFT purchase (prevented by accurate valuation) would cover the cost difference for thousands of evaluations.

\subsubsection{EQ4: Robustness and Failure Handling}

The system demonstrates strong robustness across multiple dimensions:

\textbf{Missing data}: Claude Sonnet 4.5 excels with explicit acknowledgment of unknowns and appropriate confidence reduction. GPT-5.2 performs well. Gemini 3.0 shows concerning overconfidence when data are incomplete.

\textbf{Conflicting information}: Both Claude and GPT-5.2 successfully detect conflicts and cite multiple sources with conservative recommendations. Gemini failed to identify the conflict in our test case.

\textbf{Prompt injection}: All three models successfully resist adversarial inputs embedded in user prompts or data sources. This validates that modern LLM safety training provides baseline protection against manipulation attempts.

\textbf{Policy enforcement}: While not extensively tested in this evaluation, the execution agent successfully prevented all policy violations in preliminary testing through NFTAA contract-level enforcement.

\subsection{Limitations and Threats to Validity}

\textbf{Sample size}: This evaluation examines only two scenarios with five  variations total. While these cases are representative, broader evaluation across dozens of assets and proposals would strengthen generalizability claims.

\textbf{Expert bias}: Ground truth relies on three-expert consensus, which may contain subjective bias particularly for utility assessment. Expanding the expert panel or incorporating market outcome validation would improve reliability.

\textbf{Model versions}: LLM capabilities evolve rapidly. This evaluation reflects model behavior as of December 2024. Future model versions may perform differently, though the architectural framework remains model-agnostic.

\textbf{Static evaluation}: These tests evaluate snapshot assessments. Real-world deployment would face dynamic conditions (changing prices, new information, adversarial actors at scale) not fully captured here.

\textbf{Testnet limitations}: On-chain execution testing used Sepolia and Westend testnets. Mainnet deployment would encounter additional challenges (gas price volatility, network congestion, MEV) requiring further validation.

\subsection{Discussion}

This evaluation validates the core claims of this thesis through concrete, replicable scenarios. The results demonstrate that:

\beginc{enumerate}
    \item The A/D/U value framework successfully translates theoretical concepts into practical assessment capability, achieving expert-level accuracy when implemented with Claude Sonnet 4.5
    \item intent-based weighting enables a single framework to serve diverse user goals, from speculation to long-term holding to functional use
    \item Computational efficiency makes the approach economically viable for high-volume applications, with 98-99\% cost reduction vs. manual evaluation
    \item The architecture exhibits appropriate robustness to missing data and adversarial inputs, though model selection significantly impacts reliability
\end{enumerate}

The stark performance difference between models (Claude achieving perfection, Gemini making critical errors) underscores an important finding: architectural design alone is insufficient. Model selection represents a critical system design decision that directly impacts trustworthiness.

For production deployment, these results suggest Claude Sonnet 4.5 as the preferred reasoning engine given its superior accuracy and robustness characteristics. However, the modular architecture allows component substitution as model capabilities evolve, ensuring the system can leverage future improvements without redesign.
