%Conclusion
\section{Conclusions and Future Work}
\label{sec:conclusions}

The Internet of Value is no longer a thesis-stage idea. It is the working infrastructure under which heterogeneous assets are represented, owned, transferred, and settled across treasury systems, decentralised finance, non-fungible token markets, and on-chain governance. What this infrastructure does not provide is a formal layer for interpreting what those assets are worth under a stated user or governance intent before any value-relevant action is authorised. This thesis has argued that the missing layer is a decision model, not another wallet, oracle, or voting interface, and has proposed Intent-Conditioned Value Projection (ICVP) as that model.

For RQ1 the thesis contributes a formal model in which stated intent acts as a projection $P_i$ over heterogeneous evidence spaces and an instantiation $\Phi_i$ that turns the projected evidence into intent-instantiated Availability, Durability, and Utility. The output is a valuation trace and a decision in $\{\textit{proceed}, \textit{defer}, \textit{decline}\}$ rather than a narrative explanation, which is what makes the model auditable. The conceptual content of the model is that intent does not only change how much each valuation dimension matters. It changes what counts as Availability, Durability, and Utility in the first place. The same valuation taxonomy is reused across asset classes, but its content is reset by the projection at every change of intent.

For RQ2 the thesis contributes a method that puts ICVP into practice and a precise role for the NFTAA primitive inside that method. The method decomposes the agentic pipeline into evidence acquisition, intent-specific projection and instantiation, trace construction with deterministic validation, human approval, and policy-constrained transaction preparation. NFTAA is the account-level enforcement layer for trace-gated authorisation. It does not compute value. It enforces that only human-approved, policy-compliant actions linked to a valid valuation trace can change asset state on-chain. The two formal safety properties that frame the method are stated and argued by construction. Evidence-Completeness Safety states that a confident \textit{proceed} decision implies the presence and provenance of intent-required evidence. Trace-Gated Authorisation states that authorisation of a value-relevant transaction implies a valid trace, complete intent-required evidence, a \textit{proceed} decision, a transaction that matches the trace, and a permitting policy. Neither property attempts to prove that the underlying language model is correct or that any decision is universally true.

For RQ3 the thesis contributes a framework-validation evaluation across twelve tasks that span Pudgy Penguin NFT decisions under four intents, stETH productive-asset decisions under four intents, and Polkadot OpenGov treasury proposals under public-resource stewardship. The full ICVP configuration is compared with a single-dimensional baseline that collapses each task to its dominant visible signal and a fixed-weight baseline that retains Availability, Durability, and Utility but removes intent-conditioned projection and instantiation. The full configuration agrees with the reference decisions on the full set of tasks in the reported runs, while the two baselines lose ground precisely on the tasks where the intent reshapes which evidence matters. Robustness checks under missing evidence, conflicting evidence, and prompt injection are consistent with the Evidence-Completeness Safety property in the sense that confident \textit{proceed} decisions on incomplete evidence are rare, and conflicts and injection attempts are recorded in the trace rather than silently absorbed.

The thesis claims to define a formal decision model for intent-conditioned value interpretation in the Internet of Value. It does not claim a universal theory of value, that the agentic system is correct in any stronger sense than the trace-gating allows, that experts are absolute ground truth, or that agents should autonomously control user assets. Human approval remains necessary in every value-relevant action. The agent retrieves evidence, interprets it under the stated intent, proposes a valuation trace, and prepares a transaction. The human approves and the NFTAA-level policy checks. This division of labour is what the formal proof targets are designed to make defensible.

Several directions remain open. The intent profiles used in the operational aggregation are set rather than learned, and learning these profiles from validated traces is a natural next step that would not require relaxing the safety properties. Ethical and sustainability factors are not yet first-class evidence categories under any of the studied intents and are likely to matter for medical-data, regulated-asset, and public-good settings that this thesis stops short of. The threat model assumes that adapter implementations and data sources are reliable, and a sharper treatment of adversarial data providers, compromised endpoints, and structured prompt injection embedded in retrieved content would strengthen the robustness story. Cryptographic verification of evidence provenance and consensus-based validation across multiple independent sources are concrete avenues for that work. Finally, the evaluation regimes here are deliberately compact. Extending the framework to medical data, tokenised real-world claims, and cross-chain governance is the kind of asset-class generalisation that ICVP was designed to absorb without redesign, and future work should test that ambition against richer evidence spaces and more contested reference panels.
